---
title: "Covid 19"
knit: (function(input_file, encoding) {
  out_dir <- 'docs';
  rmarkdown::render(input_file,
 encoding=encoding,
 output_file=file.path(dirname(input_file), out_dir, 'covid.html'))})
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE)
library(magrittr)
library(dplyr)
library(readr)
library(tidyr)
library(lubridate)
library(pander)
library(ggplot2)
library(scales) #commas 
library(gt) #format tables
library(gridExtra) # grid.arrange
```

## Background and Data

Although by this point some of you may be avoiding the information overload surrounding Covid-19, understanding what is currently happening by getting up-to-date data directly from the source, and extracting one's own conclusions can be empowering. 

At this point of the term you've already an R toolbox broad enough to tackle the massive amount of data found in Johns Hopkins COVID-19 repository [(click here)](https://github.com/CSSEGISandData/COVID-19/tree/master/csse_covid_19_data).  These data has been updated daily since the COVID-19 pandemic started. I also included links to other supplementary data sets to **potentially** explore the effectiveness of measures taken throughout the pandemic, including-mask use and vaccination. The datasets included are: 

1. **covid.ts.cases**: daily time series for the confirmed number of COVID-19 cases at the county level for the US (file description [here](https://github.com/CSSEGISandData/COVID-19/blob/master/csse_covid_19_data/csse_covid_19_time_series/README.md)).

2.  **covid.ts.deaths**: daily time series for the confirmed number of COVID-19 deaths at the county level for the US (file description [here](https://github.com/CSSEGISandData/COVID-19/blob/master/csse_covid_19_data/csse_covid_19_time_series/README.md)).

3. **covid.usa.daily**: COVID-19 USA daily state reports with the number of confirmed cases between April 14th and May 7th (file description [here](https://github.com/CSSEGISandData/COVID-19/tree/master/csse_covid_19_data#daily-reports-csse_covid_19_daily_reports)).


4. **Vaccination data**: state level COVID-19 daily vaccination numbers time series data from the Johns Hopkins University repository (file description [here](https://github.com/govex/COVID-19/blob/master/data_tables/vaccine_data/us_data/readme.md), )

5. **Mask use data**: information on mask use in the NY Times repository (file description [here](https://github.com/nytimes/covid-19-data/tree/master/mask-use#readme)).

6. **State policy data**: data files (one file by state) about dates and description of policies going into/out of effect.  To load data for a particular state go to [this link](https://github.com/govex/COVID-19/tree/govex_data/data_tables/policy_data/table_data/Current), find the name 
of the state file you want to work with.  

Here is the data:

```{r load_data}
# confirmed COVID-19 time series cases US (trimmed to include from 11/01/2020 to 05/12/2021)
covid.usa.ts.confirmed <- read_csv('https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_time_series/time_series_covid19_confirmed_US.csv') %>% select(c(1:11,296:488))

# Confirmed COVID-19 time series deaths US (trimmed to include from 11/01/2020 to 05/12/2021)
covid.usa.ts.deaths <- read_csv('https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_time_series/time_series_covid19_deaths_US.csv') %>% select(c(1:11,296:488))

# Daily data summary by state for 05-12-2021
covid.usa.daily <- read_csv("https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_daily_reports_us/05-12-2021.csv") 

# US vaccinated people data
vacc.people <- read_csv("https://raw.githubusercontent.com/govex/COVID-19/master/data_tables/vaccine_data/us_data/time_series/people_vaccinated_us_timeline.csv")

# mask use data
maskuse <- read_csv("https://raw.githubusercontent.com/nytimes/covid-19-data/master/mask-use/mask-use-by-county.csv")

# Texas
policytrackerTX <- read_csv("https://raw.githubusercontent.com/govex/COVID-19/govex_data/data_tables/policy_data/table_data/Current/Texas_policy.csv")

# Florida
policytrackerFL <- read_csv("https://raw.githubusercontent.com/govex/COVID-19/govex_data/data_tables/policy_data/table_data/Current/Florida_policy.csv")

# Hawaii
policytrackerHI <- read_csv("https://raw.githubusercontent.com/govex/COVID-19/govex_data/data_tables/policy_data/table_data/Current/Hawaii_policy.csv")

# Maine
policytrackerME <- read_csv("https://raw.githubusercontent.com/govex/COVID-19/govex_data/data_tables/policy_data/table_data/Current/Maine_policy.csv")

# Montana
policytrackerMT <- read_csv("https://raw.githubusercontent.com/govex/COVID-19/govex_data/data_tables/policy_data/table_data/Current/Montana_policy.csv")

# California 
policytrackerCA <- read_csv("https://raw.githubusercontent.com/govex/COVID-19/govex_data/data_tables/policy_data/table_data/Current/California_policy.csv")

```

## Part 1: Wrangling the COVID-19 time series data

1. Using the time series data sets `covid.usa.ts.confirmed` and `covid.use.ts.deaths`, which are both at the county level and are in wide format, reshape them into long format (using either of the functions `gather` or `pivot_longer`) to generate a single new data frame with the daily time series **BY STATE** including both number of confirmed cases and deaths, so that you have one row for each combination of state and date.  Call this new data frame `covid.usa.states.ts`.  **Note**: after reshaping your file into long format, your new `date` column (or however you decide to call it) needs to be converted into a `Date-Time` type variable. For example, if your variable is called `my.date.variable`, you can make this conversion using `lubridate::mdy(my.date.variable)`.

```{r, echo=F}
# long.deaths.by.state : long form of covid.usa.ts.deaths
# pivot_longer : takes in the columns from "10/31/20":"5/11/21"
# names_to : renames that column dates
# values_to : output values as "deaths"
# select : columns Province_State , dates, and deaths
# group_by : Province_State and dates
# summarize : set deaths = to the sum of deaths (per state, per day)
# mutates : formats Dates into month/day/year
long.deaths.by.state <- covid.usa.ts.deaths %>%
  pivot_longer(cols = "11/1/20":"5/11/21",
               names_to = "Dates",
               values_to = "Deaths") %>%
  select(Province_State, Dates, Deaths) %>%
  group_by(Province_State, Dates) %>%
  summarise(Deaths = sum(Deaths))  %>%
  mutate(Dates = mdy(Dates))
# long.confimed.by.state : long form of covid.usa.ts.confimed
# similar code as above but Confirmed_Cases instead of Deaths
long.confirmed.by.state <- covid.usa.ts.confirmed %>%
  pivot_longer(cols = "11/1/20":"5/11/21",
               names_to = "Dates",
               values_to = "Confirmed_Cases") %>%
  select(Province_State, Dates, Confirmed_Cases) %>%
  group_by(Province_State, Dates) %>%
  summarise(Confirmed_Cases = sum(Confirmed_Cases))  %>%
  mutate(Dates = mdy(Dates))
# covid.usa.states.ts : single data frame with daily time series by state
# select dates, deaths, and province_state from long.deaths.by.state df
# full join with long.confirmed.by.state
# select dates, confirmed_cases and province_state from long.confimed.by.state
# join by Province_State and Dates
covid.usa.states.ts <- 
  long.deaths.by.state %>%
  select(Dates, Deaths, Province_State) %>%
  full_join(long.confirmed.by.state %>%
              select(Dates, Confirmed_Cases, Province_State), 
            by = c("Province_State","Dates"))
# output new dataframe covid.usa.states.t
head(covid.usa.states.ts, n = 5)
```

2. Append to `covid.usa.states.ts` (created in previous problem) all of the information from matching rows in `vacc.people` (without repeating columns with the same info in the two data sets). 

```{r, echo=F}
# Appending covid.usa.states.ts to include people fully and partially vaccinated
# Using the same df select dates, province_state, deathes, and confirmed_cases
# full join with vacc.people
# select Date, Province_State, People_Fully_Vaccinated, and People_Partially_Vaccinated
# join by "Dates"="Date" and "Province_State"
covid.usa.states.ts <- 
  covid.usa.states.ts %>% 
  select(Dates, Province_State, Deaths, Confirmed_Cases) %>%
  full_join(vacc.people %>% 
              select(Date, Province_State, People_Fully_Vaccinated, People_Partially_Vaccinated),
            by = c("Dates"="Date","Province_State")) 
# output appended covid.usa.states.ts
head(covid.usa.states.ts, n = 5)
```

Note: People_Fully_Vaccinated and People_Partially_Vaccinated show as NA because of the dates displayed were before the vaccine was released. 

## Part 2: Let's use the data

1. Using `covid.usa.daily` select *3 highly impacted states*, *3 mildly impacted states* by COVID-19, where by 'highly impacted' I mean the states with high numbers of confirmed cases.  

```{r, echo=F, results=F}
# arrange covid.usa.daily in decending order of Confirmed cases 
covid.usa.daily %>%
  arrange(desc(Confirmed))
```

**Highly Impacted States** 

* California (3,763,281 confirmed)
* Texas (2,922,680 confirmed)
* Florida (2,278,549 confirmed)


**Mildly Impacted States** 

* Hawaii (34,917 confirmed)
* Maine (64,748 confirmed)
* Montana (110,317 confirmed)

2. Create a visualization of the evolution of confirmed cases, deaths, and people vaccinated for each of the 6 states identified as *highly* and *mildly* impacted (use the `covid.usa.states.ts` data.frame to create the figure).

```{r, echo=F}
# state vectors
highly.impacted <- c("California", "Texas", "Florida")
mildly.impacted <- c("Hawaii", "Maine", "Montana")
# column title vector
col.titles <- c("Dates", "Province_State", "People","title", "id")
# color palettes 
highly.impacted.colors <- c("royalblue1", "steelblue1","turquoise4")
mildly.impacted.colors <- c("coral", "orangered2","darkred")
```

```{r, echo=F}
# vectors 
# id's (each set covers 225 days for 3 states, 225*3 = 675)
a <- 1:684
b <- 685:1368
# state vectors
highly.impacted <- c("California", "Texas", "Florida")
mildly.impacted <- c("Hawaii", "Maine", "Montana")
# column title vector
col.titles <- c("Dates", "Province_State", "People","title", "id")
# color palettes 
highly.impacted.colors <- c("royalblue1", "steelblue1","turquoise4")
mildly.impacted.colors <- c("coral", "orangered2","darkred")
```

```{r, echo=F}
# data org. 
# note 1 : scale for deaths does not match cases and vaccinations
# note 2 : scale for mild and high do not match 
# mild cases, and vaccinations
mild.cases <- covid.usa.states.ts %>%
  filter(Province_State%in%mildly.impacted) %>%
  rename(People = Confirmed_Cases) %>%
  cbind(title = "Confirmed COVID Cases",
        id = a) 
mild.vacc <- covid.usa.states.ts %>%
  filter(Province_State%in%mildly.impacted) %>%
  rename(People = People_Fully_Vaccinated) %>%
  cbind(title = "People Fully Vaccinated",
        id = b) 
# high cases, and vaccinations 
high.cases <- covid.usa.states.ts %>%
  filter(Province_State%in%highly.impacted) %>%
  rename(People = Confirmed_Cases) %>%
  cbind(title = "Confirmed COVID Cases",
        id = a) 
high.vacc <- covid.usa.states.ts %>%
  filter(Province_State%in%highly.impacted) %>%
  rename(People = People_Fully_Vaccinated) %>%
  cbind(title = "People Fully Vaccinated",
        id = b)
# high and mild deaths 
high.deaths <- covid.usa.states.ts %>%
  filter(Province_State%in%highly.impacted) %>%
  rename(People = Deaths) %>%
  cbind(title = "COVID Deaths",
        id = a)
mild.deaths <- covid.usa.states.ts %>%
  filter(Province_State%in%mildly.impacted) %>%
  rename(People = Deaths) %>%
  cbind(title = "COVID Deaths",
        id = b)
```

```{r, echo=F} 
# joins 
# note : death scales don't match 
# mild join
mild.merge <- mild.cases %>%
  select(col.titles) %>%
  full_join(mild.vacc %>%
              select(col.titles),
            by = col.titles)
# high join
high.merge <- high.cases %>%
  select(col.titles) %>%
  full_join(high.vacc %>%
              select(col.titles),
            by = col.titles) 
```

```{r, echo=F, warning=F}
# mild cases and vaccinations 
m <- mild.merge %>% 
  ggplot(aes(x = Dates, y = People, color = Province_State)) +
  geom_line(size = 1) +
  ggtitle("Mildly Impacted States") +
  xlab("Month") + ylab("Number of People") +
  facet_wrap(~ title, nrow = 1) +
  scale_x_date(date_breaks = "1 month", date_labels = "%m") +
  scale_y_continuous(label = comma) +
  theme_minimal() + 
  theme(strip.text = element_text(face = 'bold'),
        plot.title = element_text(hjust = 0.5)) +
  scale_color_manual(name = "States",
                     values = mildly.impacted.colors)
# high cases and vaccinations
h <- high.merge %>% 
  ggplot(aes(x = Dates, y = People, color = Province_State)) +
  geom_line(size = 1) +
  ggtitle("Highly Impacted States") +
  xlab("Month") + ylab("Number of People") +
  facet_wrap(~ title, nrow = 1) +
  scale_x_date(date_breaks = "1 month", date_labels = "%m") +
  scale_y_continuous(label = comma) +
  theme_minimal() + 
  theme( strip.text = element_text(face = 'bold'),
         plot.title = element_text(hjust = 0.5)) +
  scale_color_manual(name = "States",
                     values = highly.impacted.colors)
# high deaths 
d1 <- high.deaths %>% 
  ggplot(aes(x = Dates, y = People, color = Province_State)) +
  geom_line(size = 1) +
  ggtitle("Highly Impacted States") +
  xlab("Month") + ylab("Deaths") +
  facet_wrap(~ title, nrow = 1) + # keeps subtitle 
  scale_x_date(date_breaks = "1 month", date_labels = "%m") +
  scale_y_continuous(label = comma) +
  theme_minimal() + 
  theme( strip.text = element_text(face = 'bold'),
         plot.title = element_text(hjust = 0.5)) +
  scale_color_manual(name = "States",
                     values = highly.impacted.colors)
# mild deaths 
d2 <- mild.deaths %>% 
  ggplot(aes(x = Dates, y = People, color = Province_State)) +
  geom_line(size = 1) +
  ggtitle("Mildly Impacted States") +
  xlab("Month") + ylab("Deaths") +
  facet_wrap(~ title, nrow = 1) + # keeps subtitle 
  scale_x_date(date_breaks = "1 month", date_labels = "%m") +
  scale_y_continuous(label = comma) +
  theme_minimal() + 
  theme(strip.text = element_text(face = 'bold'),
        plot.title = element_text(hjust = 0.5)) +
  scale_color_manual(name = "States",
                     values = mildly.impacted.colors)
# arranged plots  
grid.arrange(h, m, nrow = 2)
grid.arrange(d1, d2, nrow = 2)
```

3. Do you see any interesting change in the trajectories of the corresponding time series for the number of cases and deaths taking place with vaccinations?  Produce **meaningful** summaries that enable you to quantify this change (e.g., average number of new cases in windows of 90 days before vs 90 days after vaccination started). One or two summary measures is good.  Make either a table or a figure to display your findings and comment on them.

```{r, echo=F, warning=F}
# function for 90 day summaries of avg cases and deaths
state.90.day.sums <- function(state, date1, date2, title){
  covid.usa.states.ts %>%
  filter(Province_State%in%c(state)) %>%
  group_by(Dates >= date1 & Dates <= date2) %>%
  summarise(
    "Average Confirmed Cases" = mean(Confirmed_Cases, na.rm = TRUE),
    "Average Deaths" = mean(Deaths, na.rm = TRUE)) %>% 
  filter(`Dates >= date1 & Dates <= date2` == TRUE) %>%
    select("Average Confirmed Cases", "Average Deaths") %>%
    pivot_longer(cols = "Average Confirmed Cases":"Average Deaths",
                 names_to = title)}
# function that joins the before data with the after and creates a pretty table 
state.90.day.join <- function(before, after, state){
  before %>% 
    select("ninety_days_before_vacc","value") %>%
    left_join(after %>% select("ninety_days_after_vacc","value"),
              by = c("ninety_days_before_vacc"="ninety_days_after_vacc")) %>%
    rename("cases_v_deaths" = "ninety_days_before_vacc",
           "ninety_days_before_vacc" = "value.x",
           "ninety_days_after_vacc" = "value.y") %>%
    gt() %>%
    fmt_number(columns = vars(ninety_days_before_vacc), decimals = 2) %>%
    fmt_number(columns = vars(ninety_days_after_vacc), decimals = 2) %>%
  tab_header(
    title = md(state),
    subtitle = "Average Daily COVID Cases and Deaths 90 Days Before and After the Vaccine"
  ) %>%
  cols_label(
    ninety_days_before_vacc = "90 Days Before Vaccine",
    ninety_days_after_vacc  = "90 Days After Vaccine",
    cases_v_deaths = "") 
}
# used covid.usa.states.ts data to find first data with data for "fully vaccinated"
# used as.Date() +/- 90 to figure out dates 
# Hawaii 
before.HW <- state.90.day.sums("Hawaii","2020-11-11","2021-02-09", "ninety_days_before_vacc") 
after.HW <- state.90.day.sums("Hawaii","2021-02-10","2021-05-11", "ninety_days_after_vacc")
HW.join <- state.90.day.join(before.HW, after.HW, "Hawaii")
HW.join
# Maine
before.ME <- state.90.day.sums("Maine","2020-10-09","2021-01-07", "ninety_days_before_vacc") 
after.ME <- state.90.day.sums("Maine","2021-01-08","2021-04-08", "ninety_days_after_vacc")
ME.join <- state.90.day.join(before.ME, after.ME, "Maine")
ME.join
# Montana
before.MT <- state.90.day.sums("Montana","2020-10-21","2021-01-19", "ninety_days_before_vacc") 
after.MT <- state.90.day.sums("Montana","2021-01-20","2021-04-20", "ninety_days_after_vacc")
MT.join <- state.90.day.join(before.MT, after.MT, "Montana")
MT.join
# California
before.CA <- state.90.day.sums("California","2020-11-09","2021-02-07", "ninety_days_before_vacc")
after.CA <- state.90.day.sums("California","2021-02-08","2021-05-09", "ninety_days_after_vacc")
CA.join <- state.90.day.join(before.CA, after.CA, "California")
CA.join
# Florida
before.FL <- state.90.day.sums("Florida","2020-10-06","2021-01-04", "ninety_days_before_vacc")
after.FL <- state.90.day.sums("Florida","2021-01-05","2021-04-05", "ninety_days_after_vacc")
FL.join <- state.90.day.join(before.FL, after.FL, "Florida")
FL.join
# Texas 
before.TX <- state.90.day.sums("Texas","2020-10-07","2021-01-05", "ninety_days_before_vacc")
after.TX <- state.90.day.sums("Texas","2021-01-06","2021-04-06", "ninety_days_after_vacc")
TX.join <- state.90.day.join(before.TX, after.TX, "Texas")
TX.join
```

4. Using information from the `maskuse` data explore if mask use seems to correlate with your selected states being midly or highly impacted.

```{r, echo=F, warning=F}
# function for maskuse data
mask.longer <- function(state, color){
  covid.usa.ts.deaths %>%
  filter(Province_State == state) %>%
  select(FIPS,Admin2) %>%
  left_join(maskuse %>% select(COUNTYFP,NEVER,RARELY,SOMETIMES,FREQUENTLY,ALWAYS) %>%
              mutate(COUNTYFP =as.numeric(COUNTYFP)),
            by = c("FIPS"="COUNTYFP")) %>%
  pivot_longer(
    cols = "NEVER":"ALWAYS",
    names_to = "mask_use") %>%
  mutate(value = as.numeric(value)) %>%
  group_by(mask_use) %>%
  summarise(mean_mask_use = mean(value,na.rm=T)*100) %>%
  ggplot() +
  geom_col(aes(y = mask_use, x = mean_mask_use),
                 fill = color) +
  ggtitle(state) +
  xlab("Average Percent") + ylab("Frequency") +
  xlim(0,90) +
  theme_minimal() + 
  theme(
    strip.text = element_text(face = 'bold'),
    plot.title = element_text(hjust = 0.5))
}
# California 
california.mask.longer <- mask.longer("California", 'royalblue1')
# Florida
florida.mask.longer <- mask.longer("Florida", 'steelblue1')
# Texas
texas.mask.longer <- mask.longer("Texas", 'turquoise4')
# Hawaii
hawaii.mask.longer <- mask.longer("Hawaii", 'coral')
# Maine 
maine.mask.longer <- mask.longer("Maine", 'orangered2')
# Montana 
montana.mask.longer <- mask.longer("Montana", 'darkred')
grid.arrange(california.mask.longer, 
             florida.mask.longer + theme(axis.title.y = element_blank()), 
             texas.mask.longer + theme(axis.title.y = element_blank()),
             hawaii.mask.longer, 
             maine.mask.longer + theme(axis.title.y = element_blank()), 
             montana.mask.longer + theme(axis.title.y = element_blank()), 
             nrow = 2, top = "Maskuse")

```

5. Formulate and explore ONE question about the 3 highly and 3 mildly affected states with any of the datasets I have provided, or by using the policy tracker datasets corresponding to your states.

**How do open and closure days relate to number of Deaths, Confirmed Cases, and People Fully Vaccinated from 3 highly affected states (California, Florida, Texas), and 3 mildly affected states (Hawaii, Maine, Montana)?**

Note:

<span style="color: yellow;">Opening are shown with gold vertical lines</span>

<span style="color: red;">Closures are shown with red vertical lines</span>

```{r, echo=F, warning=F}
# function for confirmed cases 
cases.open.close <- function(data, state, color.var){
  #open dates
  open <- data %>% 
    filter(!is.na(`Opening (State)`)) %>%
    select(date,
         `Opening (State)`) %>%
  mutate(date = dmy(date)) 
  #closure dates
  close <- data %>%
  filter(!is.na(`Restrict/Close`)) %>%
  select(date,
         `Restrict/Close`) %>%
  mutate(date = dmy(date)) 
  #graph 
  covid.usa.states.ts %>%
  filter(Province_State%in%c(state)) %>%
  ggplot(aes(x = Dates, y = Confirmed_Cases)) +
  geom_line(size = 1, color = color.var) +
  geom_vline(xintercept = open$date, color = "gold1", size = 1) +
  geom_vline(xintercept = close$date, color = "red", size = 1) +
  labs(title = "Confirmed Cases") +
  xlab("Month") + ylab("Confirmed Cases") +
  scale_x_date(date_breaks = "1 month", date_labels = "%m") +
  scale_y_continuous(label = comma) +
  theme_minimal() +
  theme(
    strip.text = element_text(face = 'bold'),
    plot.title = element_text(hjust = 0.5)) 
}
# function for deaths
deaths.open.close <- function(data, state, color.var){
  #open dates
  open <- data %>% 
    filter(!is.na(`Opening (State)`)) %>%
    select(date,
         `Opening (State)`) %>%
  mutate(date = dmy(date)) 
  #closure dates
  close <- data %>%
  filter(!is.na(`Restrict/Close`)) %>%
  select(date,
         `Restrict/Close`) %>%
  mutate(date = dmy(date)) 
  #graph 
  covid.usa.states.ts %>%
  filter(Province_State%in%c(state)) %>%
  ggplot(aes(x = Dates, y = Deaths)) +
  geom_line(size = 1, color = color.var) +
  geom_vline(xintercept = open$date, color = "gold1", size = 1) +
  geom_vline(xintercept = close$date, color = "red", size = 1) +
  labs(title = "Deaths") +
  xlab("Month") + ylab("Deaths") +
  scale_x_date(date_breaks = "1 month", date_labels = "%m") +
  scale_y_continuous(label = comma) +
  theme_minimal() +
  theme(
    strip.text = element_text(face = 'bold'),
    plot.title = element_text(hjust = 0.5)) 
}
# function for vaccination 
vacc.open.close <- function(data, state, color.var){
  #open dates
  open <- data %>% 
    filter(!is.na(`Opening (State)`)) %>%
    select(date,
         `Opening (State)`) %>%
  mutate(date = dmy(date)) 
  #closure dates
  close <- data %>%
  filter(!is.na(`Restrict/Close`)) %>%
  select(date,
         `Restrict/Close`) %>%
  mutate(date = dmy(date)) 
  #graph 
  covid.usa.states.ts %>%
  filter(Province_State%in%c(state)) %>%
  ggplot(aes(x = Dates, y = People_Fully_Vaccinated)) +
  geom_line(size = 1, color = color.var) +
  geom_vline(xintercept = open$date, color = "gold1", size = 1) +
  geom_vline(xintercept = close$date, color = "red", size = 1) +
  labs(title = "People Fully Vaccinated") +
  xlab("Month") + ylab("People Fully Vaccinated") +
  scale_x_date(date_breaks = "1 month", date_labels = "%m") +
  scale_y_continuous(label = comma) +
  theme_minimal() +
  theme(
    strip.text = element_text(face = 'bold'),
    plot.title = element_text(hjust = 0.5)) 
}
# function that outputs all three graphs 
graphs <- function(data, state, color){
  state.cases.open.close <- cases.open.close(data, state, color)
  state.deaths.open.close <- deaths.open.close(data,state,color)
  state.vacc.open.close <- vacc.open.close(data,state, color)
  grid.arrange(state.cases.open.close, state.deaths.open.close, state.vacc.open.close, nrow=2, top = state)
  state.vacc.open.close
}
# states 
CA.graphs <- graphs(policytrackerCA,"California",'royalblue1')
FL.graphs <- graphs(policytrackerFL,"Florida",'steelblue1')
TX.graphs <- graphs(policytrackerTX,"Texas",'turquoise4')
HI.graphs <- graphs(policytrackerHI,"Hawaii",'coral')
ME.graphs <- graphs(policytrackerME,"Maine",'orangered2')
MT.graphs <- graphs(policytrackerMT,"Montana",'darkred')
```




